Tissue Enrichment Analysis for C. elegans Genomics

### 2016-08-26 tests/hgf_benchmarking/doc/tea_main_text.tex
remove wbids from marplot
because they are super annoying.

%%% Load packages
\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
% \RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
% \usepackage[utf8]{inputenc} %unicode support
% \usepackage[applemac]{inputenc} %applemac support if unicode package fails
% \usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{lmodern}% http://ctan.org/pkg/lm
\usepackage{subcaption} %for subfigures
\usepackage{graphicx} %graphics support to specify path
\usepackage{csvsimple} %insert tables using .csv files
\usepackage{caption}
\usepackage{adjustbox}

@@

\begin{abstract} % abstract
\parttitle{Background} %if any
Over the last ten years, there has been explosive development in methods
for measuring gene expression. These methods can identify thousands of altered 
genes between conditions, but understanding these datasets and forming hypotheses based on them remains challenging. One way to analyze these datasets is to associate ontologies (hierarchical, descriptive vocabularies
with controlled relations between terms) with genes and to look for enrichment of specific terms. Although gene ontology (GO) is available for   \emph{Caenorhabditis~elegans}, it does not include anatomical information.
\parttitle{Results} %if any 
We have developed a tool for identifying  enrichment of \emph{C.~elegans} tissues among gene sets and generated a website GUI where users can access this tool. Since a common drawback to ontology enrichment analyses is result verbosity, we developed a very simple trimming algorithm to reduce the ontology size by an order of magnitude. We adjusted these filters and validated our tool using a set of 30  gold standards from Expression Cluster data in WormBase and we show our tool can even discriminate between embryonic and larval tissues and can even identify tissues down to the single-cell level. We used our tool to identify multiple neuronal tissues that are down-regulated due to pathogen infection in \emph{C.~elegans}.
\parttitle{Conclusions} %if any
Our Tissue Enrichment Analysis (TEA) can be found within WormBase, and can be downloaded using Python's standard pip installer. It tests a slimmed-down \emph{C.~elegans} tissue ontology for enrichment of specific terms and provides users with a  text and graphic representation of the results. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
\section*{Background}
	RNA-seq and other high-throughput methods in biology have the ability to identify thousands of genes that are altered between conditions. These genes are often correlated in their biological characteristics or functions, but identifying these functions remains challenging. To interpret these long lists of genes, biologists need to abstract genes into concepts that are biologically relevant to form hypotheses about what is happening in the system. One such abstraction method relies on Gene Ontology (GO). GO provides a controlled set of hierarchically ordered terms~\cite{TheGeneOntologyConsortium2000a, TheGeneOntologyConsortium2015} that provide detailed descriptions about the molecular, cellular or biochemical functions of any gene. For a given gene list, existing software programs can query whether a particular term is enriched~\cite{Mi2009, McLean2010, Huang2009, Pathan2015}. One area of biological significance that GO does not include is anatomy. One way to address this shortcoming is to use a `tissue ontology' that provides a complete anatomical description for an organism (e.g `tissue', `organ' or `specific cell'), in this case for \emph{C.~elegans}. Such an ontology has been described previously for this organism~\cite{Lee2003}. Cells and tissues are physiologically relevant units with broad, relatively well-understood functionalities amenable to hypothesis formation. The \emph{C.~elegans} database, WormBase~\cite{Howe2016}, maintains a curated list of gene expression data from the literature. Here we provide a new framework that analyzes a user-input list for enrichment of specific cells and tissues.

Another problem frequently associated with GO enrichment analysis is that it is often difficult to interpret due to the large number of terms associated with a given gene. DAVID, a common tool for GO enrichment analysis, clusters enriched terms into broad categories~\cite{Huang2007}, whereas PANTHER~\cite{Mi2009, Mi2013} attempts to solve this issue by employing a manually reduced ontology, GOslim (pers. comm., H. Yu and P. Thomas). To reduce verbosity, we have filtered our ontology using a small set of well-defined criteria to remove terms that do not contribute additional information. To our knowledge, such filtering has not been performed in an algorithmic fashion for a biological ontology before---indeed, DAVID does not employ term trimming \emph{a priori} of testing, but rather fuzzy clustering \emph{post} testing to reduce the number of ontology terms. Other pruning methods do exist (see for example~\cite{Kim2007, Mambretti2013}), but the pruning is query-dependent or generates a brand new `brief ontology' which satisfies a set of logic relationships and has certain connectivity requirements. We do not propose to regenerate a new `brief ontology', but instead we use our approach to select those nodes that have sufficient annotated evidence for statistical testing. We believe our trimming methodology strikes a good balance between detailed tissue calling and conservative testing.

@@

\section*{Results}
\subsection*{Generating a Gene-Tissue Dictionary by Specific Node Selection}
\subsubsection*{Reducing term redundancy through a similarity metric}
For our tool, we employ a previously generated cell and tissue ontology for \emph{C.~elegans}\cite{Lee2003}, which is maintained and curated by WormBase. This ontology contains thousands of anatomiy terms, but not every term is equally well-annotated. As a first step to generate our tissue enrichment software, we wished to select tissue terms that were reasonably well-annotated, yet specific enough to provide insight and not redundant with other terms. For example, nematodes have a number of neurons that are placed symmetrically along the left/right body axis, and are functionally similar. These left/right neuronal pairs (which are sisters in the ontology) have almost identical annotations, with at most one or two gene differences between them, and therefore we cannot have statistical confidence in differentiating between them. As a result, testing these sister terms provides no additional information compared with testing only the parent node to these sisters. To identify redundancy, we defined two possible similarity metrics (see \emph{Methods} section and Figure~\ref{fig:simdiagram}) that can be used to identify ontology sisters that have very high similarity between them. Intuitively, a set of sisters can be considered very similar if they share most gene annotations. Within a given set of sisters, we can calculate a similarity score for a single node by counting the number of unique annotations it contains and dividing by the total number of unique annotations in the sister set. Having assigned to each sister a similarity score, we can identify the \textbf{average} similarity score for this set of sisters, and if this average value exceeds a threshold, these sisters are not considered testable candidates. An alternative method is check whether \textbf{any} of the scores exceeds a predetermined threshold, and if so remove this sister set from the ontology. We referred to these two scoring criteria as `\textbf{avg}' and `\textbf{any}' respectively.

@@

Additionally, we reasoned that, for any parent node, if all its daughters were selected for testing, there was no additional benefit to test the parent. We removed parent nodes from the analysis if all their daughter nodes passed the annotation threshold (see Figure~\ref{fig:trim_roots}). We called this a ceiling filter. Applying these three filters reduced the number of ontology terms by an order of magnitude.

@@

Once the p-values for each term have been calculated, we apply a standard FDR correction using a Benjamini-Hochberg step-up algorithm~\cite{Benjamini1995}. FDR corrected p-values are called q-values. Genes that have a q-value less than a given alpha are considered significant. Our default setting is an alpha of 0.1, which is a standard threshold broadly agreed upon by the RNA-seq community (see for example \cite{Love2014}). This threshold cannot be altered in the web GUI, but is user tunable through our command-line implementation.

@@

%dictionaries that work well
To select between dictionaries generated between 50, 33 and 25 annotation cut-offs, and also to ensure the terms that are selected as enriched by our algorithm are reasonable, we looked in detail at the enrichment analysis results. Most results were comparable and expected. For some sets, all dictionaries performed well. For example, in our `all neuron enriched sets'~\cite{Spencer2011, Watson2008a} all terms were neuron-related regardless of the dictionary used (see Table~\ref{tab:interagree}). On the other hand, for a set enriched for germline precursor expression in the embryo~\cite{Spencer2011}, the 50 cutoff dictionary was only able to identify `oocyte WBbt:006797' which is not a germline precursor, although it is germline related; whereas the two smaller dictionaries singled out actual germline precursor cells---at the 33 cutoff, our tool identified the larval germline precursor cells `Z2' and `Z3' as enriched, and at the 25 gene cutoff the embryonic germline precursor terms `P$_4$',`P$_3$' and `P$_2$' were identified in addition to `Z2' and `Z3'. 
%Differences between dictionaries
We also queried an intestine precursor set~\cite{Spencer2011}. Notably, this gene set yielded no enrichment when using the 25 cutoff dictionary, nor when using the 50 cutoff dictionary. However, the 33 cutoff dictionary identified the E lineage, which is the intestinal precursor lineage in \emph{C.~elegans}, as enriched. Both of these results capture specific aspects of \emph{C.~elegans} that are well known to developmental biologists.

@@

%Dictionary intra-agreement
We assessed the internal agreement of our tool by using independent gene sets that we expected to be enriched in the same tissues. We used two  pan-neuronal sets~\cite{Spencer2011, Watson2008a}; two PVD sets~\cite{Spencer2011, Smith2010}; and two  GABAergic sets~\cite{Spencer2011, Cinar2005}. Overall, the tool has good internal agreement. On most sets, the same terms were enriched, although order was somewhat variable (see Table~\ref{tab:interagree}), and most high-scoring terms were preserved between sets. All comparisons can be found online in our Github repository (see Availability of data and materials). Overall, the dictionary generated by a 33 gene annotation cutoff with 0.95 redundancy threshold using the `any' criterion performed best, with a good balance between specificity, verbosity and accuracy, so we selected this parameter set to generate our static dictionary. As of this publication, the testable dictionary contains 261 terms.

\subsection*{Applying the tool}
We applied our tool to the RNA-seq datasets developed by Engelmann et al.~\cite{Engelmann2011} to gain further understanding of their underlying biology. Engelmann et al. exposed young adult worms to 5 different pathogenic bacteria or fungi for 24 hours, after which mRNA was extracted from the worms for sequencing. We ran TEA on the genes Engelmann \emph{et al} identified as up- or down-regulated. Initially we noticed that genes that are down-regulated tend to be twice better annotated on average than genes that were up-regulated, suggesting that our understanding of the worm immune system is scarce, in spite of important advances made over the last decade. Up-regulated tissues, when detected, almost always included the hypodermis and excretory duct. Three out of the five samples showed enrichment of neuronal tissues or neuronal precursor tissues amongst the down-regulated genes. As an independent verification, we also performed GO analysis using PANTHER on the down-regulated genes for \emph{D.~coniospora}. These results also showed enrichment in terms associated with neurons (see Figure~\ref{fig:Dcon}). A possible explanation for this neuronal association might be that the infected worms are sick and the neurons are beginning to shut down; an alternative hypothesis would be that the worm is down-regulating specific neuronal pathways as a behavioral response against the pathogen. Indeed, several studies~\cite{Meisel2014, Zhang2005} have provided evidence that \emph{C.~elegans} uses chemosensory neurons to identify pathogens. Our results highlight the involvement of various \emph{C.~elegans} neuronal tissues in pathogen defense.

\section*{Discussion}
%TEA Discussion
We have presented a tissue enrichment analysis tool that employs a standard hypergeometric model to test the \emph{C.~elegans} tissue ontology. We use a hypergeometric function to test a user-provided gene list for enrichment of anatomical terms in \emph{C.~elegans}. Our hope is that the physiological relevance of anatomical terms will enable researchers to make hypotheses about high-dimensionality data. Specifically, we believe an enriched term may broadly suggest one of two hypotheses: if a list is enriched in a particular anatomical region, that anatomical region is affected by the experimental treatment; alternatively, the anatomical regions that are enriched reflect biologically relevant interactions between tissues. We believe the first hypothesis is a reasonable hypothesis to make in the case of whole-worm RNA-seq data for example, whereas the second hypothesis may be more plausible in cases where a researcher already knows what tissues a particular gene list came from, as may be the case in single-cell RNA-seq.

@@

%Trimming discussion
We have also presented the first, to our knowledge, ontology trimming algorithm applied to biomedical ontologies. This algorithm, which is very easy to execute, places strong limits on the number of terms selected for testing. Due to the nature of all ontologies as hierarchical, acyclical graphs with term inheritance, term annotations are correlated along any given branch. This correlation reduces the benefits of including all terms for statistical analysis: for any given term along a branch, if that term passes significance, there is a high probability that many other terms along that branch will also pass significance. If the branch is enriched by random chance, error propagation along a branch means that many more false positives will follow. Thus, a researcher might be misled by the number of terms of correlated function and assign importance to this finding; the fact that the branching structure of GO amplifies false positive signals is a powerful argument for either reducing branch length or branch intracorrelation, or both. On the other hand, if a term is actually enriched, we argue that there is little benefit to presenting the user with additional terms along that branch. Instead, a user will benefit most from testing sparsely along the tree at a suitable specificity for hypothesis formation. Related terms of the same level should only be tested when there is sufficient annotation to differentiate, with statistical confidence, whether one term is enriched above the other. Our algorithm reduces branch length by identifying and removing nodes that are insufficiently annotated and parents that are likely to include sparse information.

We endeavoured to benchmark our tool well, but our analysis cannot address problems related to spurious term enrichment. Although we were unable to determine false-positive and false-negative rates, we do not believe this should deter scientists from using our tool. Rather, we encourage researchers to use our tool as a guide, integrating evidence from multiple sources to inform the most likely hypotheses. As with any other tool based on statistical sampling, our analysis is most vulnerable to bias in the data set. For example, expression reports are negatively biased against germline expression because of the difficulties associated with expressing DNA in this tissue~\cite{Kelly1997}. As time passes, we are certain the accuracy and power of this tool will improve thanks to the continuing efforts of the  worm research community; indeed, without the community reports of tissue expression in the first place, this tool would not be possible.

@@

\section*{Competing interests}
 	The authors declare that they have no competing interests.
\section*{Author's contributions}
    DA and PWS conceived of the project; DA developed algorithm;
	RYL made intellectual contributions to the project; RYL and JC
	developed the web GUI.
\section*{Acknowledgements}
	We thank Justin Bois for his help and support. 
	We would like to acknowledge all members of the Sternberg
	lab for helpful discussion.

@@

%gaba
\begin{figure}
  % \includegraphics[width=0.95\textwidth]{GABAcomparison3.pdf}
  \captionsetup{width= 0.95\textwidth}
  \caption{
  \csentence{Independently derived gene sets show similar results 	when tested with the same dictionary.}
  \textbf{Set 1)} GABAergic gene set from Watson~\cite{Watson2008a}.
  \textbf{Set 2)} GABAergic gene set from Spencer~\cite{Spencer2011}.
  Arrowheads highlight identical terms between both analyses. All terms refer to neurons or neuronal tissues and are GABA-associated. Dictionary with cutoff: 33; threshold: 0.95; method: `any'.
   }
  \label{fig:intragree}
\end{figure}

@@

\begin{table}[p]
	\caption{
	Comparison of results for a neuronal-enriched gene set from Watson~\cite{Watson2008a}. We ran the same genelist on a dictionary with a minimum annotation cutoff of 50, similarity threshold of 0.95 and similarity method `any' versus another with a minimum annotation cutoff of 33, similarity threshold of 0.95 and similarity method `any'. In the table, columns are labeled with their significance value (Q-value) or enrichment fold change followed by a hyphen and a number which indicates which the cutoff for the dictionary that was used for testing. Not all tissues are present in either dictionary. Hyphens denote not-applicable values, which occurs when a particular tissue is not present in both dictionaries. Full table is available in Additional File 2.
	}

	% \begin{adjustbox}{width=1\textwidth}
	% 	\centering
	% 	% \csvreader{figures/dict-comparison-50-33.csv}
	%
	% 	\csvreader[tabular={ r | c | c | c | c},
	% 		    table head ={\textbf{Tissue}} & {\textbf{Q-value-33}} & {\textbf{Q-value-50}} & {\textbf{Enrichment-33}} & {\textbf{Enrichment-50}}\\\hline,
	% 		    late after line= \\,
	% 	late after last line=\\\hline
	% 	]%
	% 	{figures/dict-comparison-50-33.csv}{}
	% 	{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
	%
	% \end{adjustbox}
	\label{tab:interagree}

>>>>

%%% Load packages
\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
\usepackage[none]{hyphenat}
% \RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
% \usepackage[utf8]{inputenc} %unicode support
% \usepackage[applemac]{inputenc} %applemac support if unicode package fails
% \usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{lmodern}% http://ctan.org/pkg/lm
\usepackage{subcaption} %for subfigures
\usepackage{graphicx} %graphics support to specify path
\usepackage{csvsimple} %insert tables using .csv files
\usepackage{caption}
\usepackage{adjustbox}

@@

\begin{abstract} % abstract
\parttitle{Background} %if any
Over the last ten years, there has been explosive development in methods
for measuring gene expression. These methods can identify thousands of  
genes altered between conditions, but understanding these datasets and forming hypotheses based on them remains challenging. One way to analyze these datasets is to associate ontologies (hierarchical, descriptive vocabularies
with controlled relations between terms) with genes and to look for enrichment of specific terms. Although Gene Ontology (GO) is available for   \emph{Caenorhabditis~elegans}, it does not include anatomical information.
\parttitle{Results} %if any 
We have developed a tool for identifying  enrichment of \emph{C.~elegans} tissues among gene sets and generated a website GUI where users can access this tool. Since a common drawback to ontology enrichment analyses is its verbosity, we developed a very simple filtering algorithm to reduce the ontology size by an order of magnitude. We adjusted these filters and validated our tool using a set of 30  gold standards from Expression Cluster data in WormBase. We show our tool can even discriminate between embryonic and larval tissues and can even identify tissues down to the single-cell level. We used our tool to identify multiple neuronal tissues that are down-regulated due to pathogen infection in \emph{C.~elegans}.
\parttitle{Conclusions} %if any
Our Tissue Enrichment Analysis (TEA) can be found within WormBase, and can be downloaded using Python's standard pip installer. It tests a slimmed-down \emph{C.~elegans} tissue ontology for enrichment of specific terms and provides users with a  text and graphic representation of the results. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%
%% Background %%
\section*{Background}
	RNA-seq and other high-throughput methods in biology have the ability to identify thousands of genes that are altered between conditions. These genes are often correlated in their biological characteristics or functions, but identifying these functions remains challenging. To interpret these long lists of genes, biologists need to abstract genes into concepts that are biologically relevant to form hypotheses about what is happening in the system. One such abstraction method relies on Gene Ontology (GO). GO provides a controlled set of hierarchically ordered terms~\cite{TheGeneOntologyConsortium2000a, TheGeneOntologyConsortium2015} that provide detailed descriptions about the molecular, cellular or biochemical functions of any gene. For a given gene list, existing software programs can query whether a particular term is enriched~\cite{Mi2009, McLean2010, Huang2009, Pathan2015}. One area of biological significance that GO does not include is anatomy. One way to address this shortcoming is to use a `tissue ontology' that provides a complete anatomical description for an organism (e.g.`tissue', `organ' or `specific cell'), in this case for \emph{C.~elegans}. Such an ontology has been described previously for this organism~\cite{Lee2003}. Cells and tissues are physiologically relevant units with broad, relatively well-understood functionalities amenable to hypothesis formation. The \emph{C.~elegans} database, WormBase~\cite{Howe2016}, maintains a curated list of gene expression data from the literature. Here we provide a new framework that analyzes a user-input list for enrichment of specific cells and tissues.

Another problem frequently associated with GO enrichment analysis is that it is often difficult to interpret due to the large number of terms associated with a given gene (which we refer to as `result verbosity'). DAVID, a common tool for GO enrichment analysis, clusters enriched terms into broad categories~\cite{Huang2007}, whereas PANTHER~\cite{Mi2009, Mi2013} attempts to solve this issue by employing a manually reduced ontology, GOslim (pers. comm., H. Yu and P. Thomas). To reduce verbosity, we have filtered our ontology using a small set of well-defined criteria to remove terms that do not contribute additional information. To our knowledge, such filtering has not been performed in an algorithmic fashion for a biological ontology before; indeed, DAVID does not employ term trimming \emph{a priori} of testing, but rather fuzzy clustering \emph{post} testing to reduce the number of ontology terms. Other pruning methods do exist (see for example~\cite{Kim2007, Garrido2012}), but the pruning is query-dependent or generates a brand new `brief ontology' which satisfies a set of logic relationships and has certain connectivity requirements. We do not propose to regenerate a new `brief ontology', but instead we use our approach to select those nodes that have sufficient annotated evidence for statistical testing. We believe our trimming methodology strikes a good balance between detailed tissue calling and conservative testing.

@@

\section*{Results}
\subsection*{Generating a Gene-Tissue Dictionary by Specific Node Selection}
\subsubsection*{Reducing term redundancy through a similarity metric}
For our tool, we employ a previously generated cell and tissue ontology for \emph{C.~elegans}~\cite{Lee2003}, which is maintained and curated by WormBase. This ontology contains thousands of anatomiy terms, but not every term is equally well-annotated. As a first step to generate our tissue enrichment software, we wished to select tissue terms that were reasonably well-annotated, yet specific enough to provide insight and not redundant with other terms. For example, nematodes have a number of neurons that are placed symmetrically along the left/right body axis, and are functionally similar. These left/right neuronal pairs (which are sisters in the ontology) have almost identical annotations, with at most one or two gene differences between them, and therefore we cannot have statistical confidence in differentiating between them. As a result, testing these sister terms provides no additional information compared with testing only the parent node to these sisters. To identify redundancy, we defined two possible similarity metrics (see \emph{Methods} section and Figure~\ref{fig:simdiagram}) that can be used to identify ontology sisters that have very high similarity between them. Intuitively, a set of sisters can be considered very similar if they share most gene annotations. Within a given set of sisters, we can calculate a similarity score for a single node by counting the number of unique annotations it contains and dividing by the total number of unique annotations in the sister set. Having assigned to each sister a similarity score, we can identify the \textbf{average} similarity score for this set of sisters, and if this average value exceeds a threshold, these sisters are not considered testable candidates. An alternative method is check whether \textbf{any} of the scores exceeds a predetermined threshold, and if so remove this sister set from the ontology. We referred to these two scoring criteria as `\textbf{avg}' and `\textbf{any}' respectively.

@@

Additionally, we reasoned that for any parent node if all its daughters were selected for testing, there was no additional benefit to test the parent. We removed parent nodes from the analysis if all their daughter nodes passed the annotation threshold (see Figure~\ref{fig:trim_roots}). We called this a ceiling filter. Applying these three filters reduced the number of ontology terms by an order of magnitude.

@@

Once the p-values for each term have been calculated, we apply a standard FDR correction using a Benjamini-Hochberg step-up algorithm~\cite{Benjamini1995}. FDR corrected p-values are called q-values. Genes that have a q-value less than a given alpha are considered significant. Our default setting is an alpha of 0.1, which is a standard threshold broadly agreed upon by the scientific community (see for example~\cite{Love2014, Pawitan2005, Storey2003}). This threshold cannot be altered in the web GUI, but is user tunable through our command-line implementation.

@@

%dictionaries that work well
To select between dictionaries generated between 50, 33 and 25 annotation cut-offs, and also to ensure the terms that are selected as enriched by our algorithm are reasonable, we looked in detail at the enrichment analysis results. Most results were comparable and expected. For some sets, all dictionaries performed well. For example, in our `all neuron enriched sets'~\cite{Spencer2011, Watson2008a} all terms were neuron-related regardless of the dictionary used (see Table~\ref{tab:intraagree}). On the other hand, for a set enriched for germline precursor expression in the embryo~\cite{Spencer2011}, the 50 cutoff dictionary was only able to identify `oocyte WBbt:006797', which is not a germline precursor although it is germline related; whereas the two smaller dictionaries singled out actual germline precursor cells---at the 33 cutoff, our tool identified the larval germline precursor cells `Z2' and `Z3' as enriched, and at the 25 gene cutoff the embryonic germline precursor terms `P$_4$',`P$_3$' and `P$_2$' were identified in addition to `Z2' and `Z3'. 
%Differences between dictionaries
We also queried an intestine precursor set~\cite{Spencer2011}. Notably, this gene set yielded no enrichment when using the 25 cutoff dictionary, nor when using the 50 cutoff dictionary. However, the 33 cutoff dictionary identified the E lineage, which is the intestinal precursor lineage in \emph{C.~elegans}, as enriched. Both of these results capture specific aspects of \emph{C.~elegans} that are well known to developmental biologists.

@@

%Dictionary intra-agreement
We assessed the internal agreement of our tool by using independent gene sets that we expected to be enriched in the same tissues. We used two  pan-neuronal sets~\cite{Spencer2011, Watson2008a}; two PVD sets~\cite{Spencer2011, Smith2010}; and two  GABAergic sets~\cite{Spencer2011, Cinar2005}. Overall, the tool has good internal agreement. On most sets, the same terms were enriched, although order was somewhat variable (see Table~\ref{fig:interagree}), and most high-scoring terms were preserved between sets. All comparisons can be found online in our Github repository (see Availability of data and materials). Overall, the dictionary generated by a 33 gene annotation cutoff with 0.95 redundancy threshold using the `any' criterion performed best, with a good balance between specificity, verbosity and accuracy, so we selected this parameter set to generate our static dictionary. As of this publication, the testable dictionary contains 261 terms.

\subsection*{Applying the tool}
We applied our tool to the RNA-seq datasets developed by Engelmann et al.~\cite{Engelmann2011} to gain further understanding of their underlying biology. Engelmann et al. exposed young adult worms to 5 different pathogenic bacteria or fungi for 24 hours, after which mRNA was extracted from the worms for sequencing. We ran TEA on the genes Engelmann \emph{et al} identified as up- or down-regulated. Initially we noticed that genes that are down-regulated tend to be twice as better annotated on average than genes that were up-regulated, suggesting that our understanding of the worm immune system is scarce, in spite of important advances made over the last decade. Up-regulated tissues, when detected, almost always included the hypodermis and excretory duct. Three of the five samples showed enrichment of neuronal tissues or neuronal precursor tissues among the down-regulated genes. As an independent verification, we also performed GO analysis using PANTHER on the down-regulated genes for \emph{D.~coniospora}. These results also showed enrichment in terms associated with neurons (see Figure~\ref{fig:Dcon}). A possible explanation for this neuronal association might be that the infected worms are sick and the neurons are beginning to shut down; an alternative hypothesis would be that the worm is down-regulating specific neuronal pathways as a behavioral response against the pathogen. Indeed, several studies~\cite{Meisel2014, Zhang2005} have provided evidence that \emph{C.~elegans} uses chemosensory neurons to identify pathogens. Our results highlight the involvement of various \emph{C.~elegans} neuronal tissues in pathogen defense.

\section*{Discussion}
%TEA Discussion
We have presented a tissue enrichment analysis tool that employs a standard hypergeometric model to test the \emph{C.~elegans} tissue ontology. We use a hypergeometric function to test a user-provided gene list for enrichment of anatomical terms in \emph{C.~elegans}. Our hope is that the physiological relevance of anatomical terms will enable researchers to make hypotheses about high-dimensionality data. Specifically, we believe an enriched term may broadly suggest one of two hypotheses: if a list is enriched in a particular anatomical region, that anatomical region is affected by the experimental treatment; alternatively, the anatomical regions that are enriched reflect biologically relevant interactions between tissues. We believe the first hypothesis is a reasonable one to make in the case of whole-worm RNA-seq data for example, whereas the second hypothesis may be more plausible in cases where a researcher already knows what tissues a particular gene list came from, as may be the case in single-cell RNA-seq.

@@

%Trimming discussion
We have also presented the first, to our knowledge, ontology term filtering algorithm applied to biomedical ontologies. This algorithm, which is very easy to execute, identifies terms that have specificity and statistical power for hypothesis testing. Due to the nature of all ontologies as hierarchical, acyclical graphs with term inheritance, term annotations are correlated along any given branch. This correlation reduces the benefits of including all terms for statistical analysis: for any given term along a branch, if that term passes significance, there is a high probability that many other terms along that branch will also pass significance. If the branch is enriched by random chance, error propagation along a branch means that many more false positives will follow. Thus, a researcher might be misled by the number of terms of correlated function and assign importance to this finding; the fact that the branching structure of GO amplifies false positive signals is a powerful argument for either reducing branch length or branch intracorrelation, or both. On the other hand, if a term is actually enriched, we argue that there is little benefit to presenting the user with additional terms along that branch. Instead, a user will benefit most from testing sparsely along the tree at a suitable specificity for hypothesis formation. Related terms of the same level should only be tested when there is sufficient annotation to differentiate, with statistical confidence, whether one term is enriched above the other. Our algorithm reduces branch length by identifying and removing nodes that are insufficiently annotated and parents that are likely to include sparse information.

We endeavoured to benchmark our tool well, but our analysis cannot address problems related to spurious term enrichment. Although we were unable to determine false-positive and false-negative rates, we do not believe this should deter scientists from using our tool. Rather, we encourage researchers to use our tool as a guide, integrating evidence from multiple sources to inform the most likely hypotheses. As with any other tool based on statistical sampling, our analysis is most vulnerable to bias in the data set. For example, expression reports are negatively biased against germline expression because of the difficulties associated with expressing transgenes in this tissue~\cite{Kelly1997}. As time passes, we are certain the accuracy and power of this tool will improve thanks to the continuing efforts of the  worm research community; indeed, without the community reports of tissue expression in the first place, this tool would not be possible.

\section*{Conclusions}

We have built a tissue enrichment tool that employs a tissue ontology previously developed by WormBase. We use a simple algorithm to identify the best ontology terms for statistical testing and in this way minimize multiple testing problems. Our tool is available within WormBase or can be downloaded for offline use via `pip install'.

@@

\section*{Competing interests}
 	The authors declare that they have no competing interests.
\section*{Author's contributions}
    DA and PWS conceived of the project; DA developed algorithm;
	RYL made intellectual contributions to the project; RYL and JC
	developed the web GUI. 
\section*{Acknowledgements}
	We thank Justin Bois for his help and support. 
	We would like to acknowledge all members of the Sternberg
	lab for helpful discussion. 
\section*{Funding}
	This work was supported by HG002223 from the National Human Genome Research 	Institute to PWS., and H.H.M.I., with which PWS is an investigator. The 	funding agencies had no input into the design, execution or interpretation 	of our experiments, nor into the writing of this manuscript.

@@

%gaba
\begin{figure}
  % \includegraphics[width=0.95\textwidth]{GABAcomparison3.pdf}
  \captionsetup{width= 0.95\textwidth}
  \caption{
  \csentence{Independently derived gene sets show similar results 	when tested with the same dictionary.}
  \textbf{Set 1)} GABAergic gene set from Watson~\cite{Watson2008a}.
  \textbf{Set 2)} GABAergic gene set from Spencer~\cite{Spencer2011}.
  Arrowheads highlight identical terms between both analyses. All terms refer to neurons or neuronal tissues and are GABA-associated. Dictionary with cutoff: 33; threshold: 0.95; method: `any'.
   }
  \label{fig:interagree}
\end{figure}

@@

\begin{table}[p]
	\caption{
	Comparison of results for a GABAergic neuronal-enriched gene set from Watson~\cite{Watson2008a} showing that results are similar regardless of annotation cutoff. We ran the same gene list on a dictionary with a minimum annotation cutoff of 50, similarity threshold of 0.95 and similarity method `any' versus another with a minimum annotation cutoff of 33, similarity threshold of 0.95 and similarity method `any'. In the table, columns are labeled with their significance value (Q-value) or enrichment fold change followed by a hyphen and a number which indicates which the cutoff for the dictionary that was used for testing. Not all tissues are present in either dictionary. Hyphens denote not-applicable values, which occurs when a particular tissue is not present in both dictionaries.
	}

	% \begin{adjustbox}{width=1\textwidth}
	% 	\centering
	% 	% \csvreader{figures/dict-comparison-50-33.csv}
	%
	% 	\csvreader[tabular={ r | c | c | c | c},
	% 		    table head ={\textbf{Tissue}} & {\textbf{Q-value-33}} & {\textbf{Q-value-50}} & {\textbf{Enrichment-33}} & {\textbf{Enrichment-50}}\\\hline,
	% 		    late after line= \\,
	% 	late after last line=\\\hline
	% 	]%
	% 	{figures/dict-comparison-50-33.csv}{}
	% 	{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
	%
	% \end{adjustbox}
	\label{tab:intraagree}
